Data Specialization for a Staged Lambda Calculus

### 2015-02-26 paper/relatedwork.tex
minor edits

Experience with binding time analysis showed that it can be difficult
to control, leading to programs whose performance can be difficult to
predict. This led to investigations based on type systems for making
explicit in the program the stage of each
computation~\cite{GJ91-lambda,NN92-twolevel} and writing ``meta
programs'' that when evaluated at a stage, give another program to be
evaluated at the next stage. Jorring and Scherlis describe this idea
meta programming in their paper (~\cite{JS86-staging}) as one of the
three possible approaches to taking advantage of
staging. Davies~\cite{Davies96} presented a logical construction of
binding-time type systems by deriving a type system via the
Curry-Howard isomorphism applied to temporal logic.  Davies and
Pfenning proposed a new type system for staged computation based on a
particular fragment of modal logic~\cite{DP01-modal}. The work on
MetaML extended type-based techniques to a full-scale language by
developing statically typed programming language based on ML that
enables the programmer to express programs with multiple
stages~\cite{Taha97,taha-thesis-99}.  MetaML's type system is similar
to Davies~\cite{Davies96} but extends it in several important ways.
Nanevksi and Pfenning further extended the these techniques by
allowing free variable to occur within staged
computations~\cite{NP05-nn}.

>>>>

Experience with binding time analysis showed that it can be difficult
to control, leading to programs whose performance can be difficult to
predict. This led to investigations based on type systems for making
explicit in the program the stage of each
computation~\cite{GJ91-lambda,NN92-twolevel} and writing ``meta
programs'' that when evaluated at a stage, give another program to be
evaluated at the next stage. Jorring and Scherlis describe this idea
meta programming in their paper (~\cite{JS86-staging}) as one of the
three possible approaches to taking advantage of
staging. Davies~\cite{davies96} presented a logical construction of
binding-time type systems by deriving a type system via the
Curry-Howard isomorphism applied to temporal logic.  Davies and
Pfenning proposed a new type system for staged computation based on a
particular fragment of modal logic~\cite{DP01-modal}. The work on
MetaML extended type-based techniques to a full-scale language by
developing statically typed programming language based on ML that
enables the programmer to express programs with multiple
stages~\cite{Taha97,taha-thesis-99}.  MetaML's type system is similar
to Davies~\cite{davies96} but extends it in several important ways.
Nanevksi and Pfenning further extended the these techniques by
allowing free variable to occur within staged
computations~\cite{NP05-nn}.
